Software Architectures to Integrate Workflow Engines in Science Gateways

### 2016-04-26 paper.tex
overall minor edits and comments

\paragraph{Specific features} Meta-workflow are not supported by default
(\texttt{O$_1$}=1). Fine-grained debugging information is usually easy
to obtain since the workflow service provides direct access to the
engine (\texttt{O$_2$}=0).

@@

\paragraph{Integration effort} Setting up a nested workflow architecture with
service invocation requires a science gateway, 2 workflow services and
2 infrastructures (\texttt{I$_1$=5}). The architecture involves 11
interactions (\texttt{I$_2$=11}): \texttt{a$_1$} (appears twice: once for each
infrastructure), \texttt{a$_2$}, \texttt{b} (twice: once for each
infrastructure), \texttt{c$_1$} (twice: once for each infrastructure),
\texttt{c$_2$} (twice: once for each workflow engine and
infrastructure) and \texttt{d} (twice: once for each workflow engine).

@@

Differences in other specific features should not be neglected. Nested
workflows (with service invocation) and workflow import (with service
invocation) are the only architectures that support meta-workflows,
which may be interesting in some cases. The availability of
fine-grained debugging information may also be critical to efficient
user support.

Table~\ref{table:overall} provides an overall comparison between
architectures, based on the metrics in Table~\ref{table:evaluation}.
Overall, sub-tasking, and workflow import (with service invocation)
perform a bit better than tight integration, service invocation and
pool. Nested workflows stands a bit behind for the reasons mentioned
previously.
\begin{table*}
\centering
\begin{tabular}{rcccccc}
                                    & \textbf{Tight}
                                    & \textbf{Service}
                                    & \textbf{Sub-task}
                                    & \textbf{Pool}
                                    & \textbf{Nested}
                                    & \textbf{Import} \\
  Integration effort (normalized total) &
                                    \cellcolor[HTML]{99FF99}{0.00}
                                    & \cellcolor[HTML]{99E899}{0.22}
                                    & \cellcolor[HTML]{99FF99}{0.00}
                                    & \cellcolor[HTML]{99D199}{0.44}
                                    & \cellcolor[HTML]{999999}{1.00}
                                      & \cellcolor[HTML]{99DD99}{0.33}\\
Robustness (normalized total) &
                                \cellcolor[HTML]{99F099}{0.14}
                                    & \cellcolor[HTML]{99D399}{0.43}
                                    & \cellcolor[HTML]{99FF99}{0.00}
                                    & \cellcolor[HTML]{99B699}{0.71}
                                    & \cellcolor[HTML]{999999}{1.00}
                                    & \cellcolor[HTML]{99D399}{0.43}\\
  Extensibility (normalized total)  & \cellcolor[HTML]{99D199}{0.44}
                                     & \cellcolor[HTML]{99D199}{0.44}
                                     & \cellcolor[HTML]{99FF99}{0.00}
                                     & \cellcolor[HTML]{99E899}{0.22}
                                     & \cellcolor[HTML]{999999}{1.00}
                                     & \cellcolor[HTML]{99E899}{0.22}\\
Scalability (normalized total)  & \cellcolor[HTML]{999999}{1.00}
                                     & \cellcolor[HTML]{99CC99}{0.50}
                                     & \cellcolor[HTML]{99CC99}{0.50}
                                     & \cellcolor[HTML]{99FF99}{0.00}
                                     & \cellcolor[HTML]{99CC99}{0.50}
                                     & \cellcolor[HTML]{99CC99}{0.50}\\
Specific features (normalized total) & \cellcolor[HTML]{99CC99}{0.50}
                                     & \cellcolor[HTML]{99CC99}{0.50}
                                     & \cellcolor[HTML]{999999}{1.00}
                                     & \cellcolor[HTML]{99CC99}{0.50}
                                     & \cellcolor[HTML]{99CC99}{0.50}
                                     & \cellcolor[HTML]{99DD99}{0.00}\\
                                    & \cellcolor[HTML]{99E699}\textbf{2.1}
                                    & \cellcolor[HTML]{99E699}\textbf{2.1}
                                    & \cellcolor[HTML]{99FE99}\textbf{1.5}
                                    & \cellcolor[HTML]{99EE99}\textbf{1.9}
                                    & \cellcolor[HTML]{999999}\textbf{4.0}
                                    & \cellcolor[HTML]{99FF99}\textbf{1.5}\\
\end{tabular}
\caption{Overall evaluation. Brighter colors and lower scores indicate better performance. Scores
  are obtained by summing the normalized total scores ($m'$ values) of
  each criterion in Table~\ref{table:evaluation} and the colors are obtained by normalizing these scores
  as done in Table~\ref{table:evaluation}. }
\label{table:overall}
\end{table*}

@@

% Summary of contributions
We reviewed architectures used to integrate workflow engines in
science gateways, we described them in a system-independent framework
suitable for comparison, we illustrated them on real systems, we
developed quantitative metrics for evaluation, we evaluated these
metrics for all the architectures, and we discussed the pros and cons
of all the architectures based on these metrics.

>>>>

\paragraph{Specific features} Meta-workflows are not supported by default
(\texttt{O$_1$}=1). Fine-grained debugging information is usually easy
to obtain since the workflow service provides direct access to the
engine (\texttt{O$_2$}=0).

@@

\paragraph{Integration effort} Setting up a nested workflow architecture with
service invocation requires a science gateway, 2 workflow services and
2 infrastructures (\texttt{I$_1$=5}). The architecture involves 11
interactions (\texttt{I$_2$=11}): \texttt{a$_1$} (appears twice: once for each
infrastructure), \texttt{a$_2$}, \texttt{b} (twice: once for each
infrastructure), \texttt{c$_1$} (twice: once for each infrastructure),
\texttt{c$_2$} (twice: once for each workflow engine and
infrastructure) and \texttt{d} (twice: once for each workflow engine).
\note{Rafael}{I like the idea of what has been discussed in the `integration effort', however I think the name may lead to misinterpretations. For example, according to Table 2, it requires more effort to integrate in a Tight architecture than in a service-based architecture. Considering the principles of software engineer, it is simpler to integrate/maintain a system using decoupled services. Maybe it should be renamed to something related to 'dependencies'.}

@@

Differences in other specific features should not be neglected. Nested
workflows (with service invocation) and workflow import (with service
invocation) are the only architectures that support meta-workflows,
which may be interesting in some cases.\note{Rafael}{I would suggest to provide some
more examples when this could be interesting.}
The availability of fine-grained debugging information may also be critical to efficient
user support.

Table~\ref{table:overall} provides an overall comparison between
architectures, based on the metrics in Table~\ref{table:evaluation}.
Overall, sub-tasking, and workflow import (with service invocation)
perform a bit better than tight integration, service invocation and
pool. Nested workflows stands a bit behind for the reasons mentioned
previously.\note{Rafael}{I would emphasize that this analysis is only for illustration purposes, since it does not account for different weights in metrics.}
\begin{table*}
\centering
\begin{tabular}{rcccccc}
                                    & \textbf{Tight}
                                    & \textbf{Service}
                                    & \textbf{Sub-task}
                                    & \textbf{Pool}
                                    & \textbf{Nested}
                                    & \textbf{Import} \\
  Integration effort (normalized total) &
                                    \cellcolor[HTML]{99FF99}{0.00}
                                    & \cellcolor[HTML]{99E899}{0.22}
                                    & \cellcolor[HTML]{99FF99}{0.00}
                                    & \cellcolor[HTML]{99D199}{0.44}
                                    & \cellcolor[HTML]{999999}{1.00}
                                      & \cellcolor[HTML]{99DD99}{0.33}\\
Robustness (normalized total) &
                                \cellcolor[HTML]{99F099}{0.14}
                                    & \cellcolor[HTML]{99D399}{0.43}
                                    & \cellcolor[HTML]{99FF99}{0.00}
                                    & \cellcolor[HTML]{99B699}{0.71}
                                    & \cellcolor[HTML]{999999}{1.00}
                                    & \cellcolor[HTML]{99D399}{0.43}\\
  Extensibility (normalized total)  & \cellcolor[HTML]{99D199}{0.44}
                                     & \cellcolor[HTML]{99D199}{0.44}
                                     & \cellcolor[HTML]{99FF99}{0.00}
                                     & \cellcolor[HTML]{99E899}{0.22}
                                     & \cellcolor[HTML]{999999}{1.00}
                                     & \cellcolor[HTML]{99E899}{0.22}\\
Scalability (normalized total)  & \cellcolor[HTML]{999999}{1.00}
                                     & \cellcolor[HTML]{99CC99}{0.50}
                                     & \cellcolor[HTML]{99CC99}{0.50}
                                     & \cellcolor[HTML]{99FF99}{0.00}
                                     & \cellcolor[HTML]{99CC99}{0.50}
                                     & \cellcolor[HTML]{99CC99}{0.50}\\
Specific features (normalized total) & \cellcolor[HTML]{99CC99}{0.50}
                                     & \cellcolor[HTML]{99CC99}{0.50}
                                     & \cellcolor[HTML]{999999}{1.00}
                                     & \cellcolor[HTML]{99CC99}{0.50}
                                     & \cellcolor[HTML]{99CC99}{0.50}
                                     & \cellcolor[HTML]{99DD99}{0.00}\\
                                    & \cellcolor[HTML]{99E699}\textbf{2.1}
                                    & \cellcolor[HTML]{99E699}\textbf{2.1}
                                    & \cellcolor[HTML]{99FE99}\textbf{1.5}
                                    & \cellcolor[HTML]{99EE99}\textbf{1.9}
                                    & \cellcolor[HTML]{999999}\textbf{4.0}
                                    & \cellcolor[HTML]{99FF99}\textbf{1.5}\\
\end{tabular}
\caption{Overall evaluation. Brighter colors and lower scores indicate better performance. Scores
  are obtained by summing the normalized total scores ($m'$ values) of
  each criterion in Table~\ref{table:evaluation} and the colors are obtained by normalizing these scores
  as done in Table~\ref{table:evaluation}. }
\label{table:overall}
\end{table*}

@@

% Summary of contributions
We reviewed architectures used to integrate workflow engines in
science gateways, we described them in a system-independent framework
suitable for comparison, we illustrated them on real systems, we
developed quantitative metrics for evaluation, we evaluated these
metrics for all the architectures, and we discussed the pros and cons
of all the architectures based on these metrics.
\note{Rafael}{One thing that is missing in the conclusion or the limitations section is to state that this work does not intend to quantify the amount of work required to develop science gateways using a specific architecture, but the goal is to highlight what are the main components and interactions for each architecture, and the main advantages and drawbacks. This should also be stated in the introduction.}
